As expected the standard RLS reaches an optimal solution the fastest.
It also reaches the optimal value for every instance.
The RLS-R variants need more iterations to find an optimal solution.
By looking at the average values more closely it seems like the average number of steps is roughly $25,000 + 70,000k \pm 5,000$.
The standard RLS is equivalent to RLS-R or RLS-N with $k=1$.
So the value of $k=1$ seems to be optimal for the RLS variants too.
The RLS-N variants on the other hand do not reach one of the two optimal solutions in any run.
This is most likely caused by their very low possibility of flipping only one bit in a single step.
They would eventually reach the optimal solution as well, but this would take to long.
The probability of only flipping one bit in a step is $\mathcal{O}(n^{1-k})$ which results in a single bit flip every $\mathcal{O}(n^{k-1})$ steps in expectation.
Because the fitness can only improve for OneMax making steps flipping more bits does not harm the fitness.
The bound for OneMax is $\mathcal{O}(nlogn)$ and with the previous result the expected number of steps is bounded by
$\mathcal{O}(n\cdot\mathcal{O}(n^{k-1})\cdot \log(n\cdot\mathcal{O}(n^{k-1}))) 
=\mathcal{O}(n^{k-1+1}\cdot (k-1+1)\cdot\log(n))
=\mathcal{O}(kn^{k}\cdot\log(n))$
This problem is not equivalent to OneMax, as a flip of the bit with the highest value inverts the fitness function to ZeroMax but the result might still hold as the bound for the standard RLS for this input is the same as for the RLS on OneMax.